# Simulations with systematicsurvey

```{r, include  = TRUE, message=FALSE, warning=FALSE}
library(ggplot2)
library(INLA)
library(raster)
library(rgeos)
library(systematicsurvey)
# can install it using 
## devtools::install_github("Craig44/systematicsurvey")

set.seed(123)
## Assume a square domain
survey_xlim = c(0,100)
survey_ylim = c(0,100)
# simualted population size
N = 5e5 
## spatial distribution
beta1 = c(4,1)
beta2 = c(1,4)

survey_area = diff(survey_ylim) * diff(survey_xlim) # Area of survey
## create a polygon
s_poly = Polygon(cbind(x = c(survey_xlim[1],survey_xlim[1],survey_xlim[2],survey_xlim[2]), y = c(survey_ylim[1],survey_ylim[2],survey_ylim[2],survey_ylim[1])))
s_polys = Polygons(list(s_poly),1)
sp_survey_poly = SpatialPolygons(list(s_polys)) ## add projection coordinater here if you are using
survey_area = rgeos::gArea(sp_survey_poly)
survey_polygon = sp_survey_poly

## sampling unit
quad_width = 0.5
quad_height = 0.5
quad_x_spacing = 10 ## spacings between quadrant/transect midpoints
quad_y_spacing = 10 ## spacings between quadrant/transect midpoints
sampling_unit_area = quad_width * quad_height
boxlet_diameter = quad_width ## will define how many boxlets there are, should be less than quadrant size
kappa = survey_area / sampling_unit_area ## sampling fraction
## variables that define the toal number of sampling units
n_col = max(survey_xlim) /  quad_x_spacing # r
n_row = max(survey_ylim) /  quad_y_spacing # r
n_strata = n_row * n_col
n_l = quad_y_spacing / quad_height# l
n_k = quad_x_spacing / quad_width # k
n_quads = n_l * n_row * n_col * n_k
within_rows = rep(1:n_l, n_k)
within_cols = sort(rep(1:n_k, n_l))
## lower and upper limits for discrete non-overlapping quadrats
quad_x = seq(0,max(survey_xlim), by = quad_width)
quad_y = seq(0,max(survey_ylim), by = quad_height)

# create a raster layer
proj_raster <- raster(resolution = quad_width, xmn = survey_xlim[1], xmx = survey_xlim[2], ymn = survey_ylim[1], ymx = survey_ylim[2])
proj_poly <- rasterToPolygons(proj_raster, dissolve=F)

## simulate spatial distribution
x_coord = rbeta(N, beta1[1], beta1[2]) * max(survey_xlim)
y_coord = rbeta(N, beta2[1], beta2[2]) * max(survey_ylim)
## rasterize
pop_raster <- rasterize(x = cbind(x_coord, y_coord), y = proj_raster,  field = 1, fun = "sum", na.rm = T)
pop_raster_pts <- rasterToPoints(pop_raster, spatial = TRUE)
# Then to a 'conventional' dataframe
pop_df  <- data.frame(pop_raster_pts)
ggplot() +
 geom_tile(data = pop_df , aes(x = x, y = y, fill = layer)) + 
    ggtitle("Population") +
  labs(x = "", y = "", fill = "Abundance")
```

Now simulate a two-dimensional systematic survey with a single PSU.

```{r sampling_unit, include  = TRUE}
PSU = sample(1:length(within_cols), size = 1)
# convert cols and rows to survey units
within_x = within_cols[PSU] * quad_width
within_y = within_rows[PSU] * quad_height

upper_x1 = seq(from = within_x, to = survey_xlim[2], by = quad_x_spacing)
upper_y1 = seq(from = within_y, to = survey_ylim[2], by = quad_y_spacing)
upper_x = rep(upper_x1, length(upper_y1))
upper_y = sort(rep(upper_y1, length(upper_x1)))
sample_y_coord = upper_y - quad_height / 2
sample_x_coord = upper_x - quad_width / 2
## data frame
Data = data.frame(x_mid = sample_x_coord, y_mid = sample_y_coord)
Data$y_i = NA
counter = 1;
## population in sampling units
for(i in 1:length(upper_x)) {
  Data$y_i[i] = sum(x_coord >= (upper_x[i] - quad_width) & x_coord <= upper_x[i] 
                      & y_coord >= (upper_y[i]  - quad_height) & y_coord <= upper_y[i])
}
Data$area = quad_height * quad_height
## convert to densities
Data$d_i = Data$y_i / Data$area

## convert it to spatial data frame
coordinates( Data ) <- ~ x_mid + y_mid
## can also add projection of data if working lat longs etc
# proj4string(Data) <- CRS("+init=epsg:4326")

## calculate sample raster
sample_raster <- rasterize(x = cbind(sample_x_coord, sample_y_coord), y = proj_raster,  field = 1, fun = "sum", na.rm = T)
## calculate polygons
samp_poly <- rasterToPolygons(sample_raster, dissolve=TRUE)
proj_indicator = apply(gContains(samp_poly,proj_poly, byid = TRUE), 1, any)
```

## SRS estimator
```{r r_srs, include  = TRUE}

srs_var_hat  = kappa^2 * var(Data@data$d_i) / (nrow(Data@data))
## standard error
sqrt(srs_var_hat)
```
## Correlation adjusted estimator
```{r r_cor_adj, include  = TRUE}

```
## Boxlet estimator

The first thing we need to do is calculate the fine lattice within $\mathcal{D}$ (these lattice points correspond to the centroids of the so-called boxlets). This is done using the `?BoxletEstimatorSamplingFrame` function. This function will return a list containing a fine lattice for all the boxlets. The other function that will calculate the boxlet variance is `?BoxletEstimator`. These are applied in the following code chunk.

```{r r_boxlet, include  = TRUE}
## build lattice that we will evaluate the variance over
## for large survey areas can take a few seconds to a minute
boxlet_frame_single = BoxletEstimatorSamplingFrame(survey_polygon, quad_width = quad_width,
                                     quad_height = quad_height, quad_x_spacing = quad_x_spacing,
                                     quad_y_spacing = quad_y_spacing, boxlet_per_sample_width = 1, 
                                     boxlet_per_sample_height = 1, trace = F)

## use the 
boxlet_estimator_single = BoxletEstimator(spatial_df = Data, survey_polygon = survey_polygon,
                                          quad_width = quad_width, quad_height = quad_height, 
                                          quad_x_spacing = quad_x_spacing, quad_y_spacing = quad_y_spacing, 
                                          boxlet_sampling_frame = boxlet_frame_single)
names(boxlet_estimator_single)
## standard error
sqrt(boxlet_estimator_single$var_total_boxlet)
```
